Comparison of two object-oriented languages: C++ and Ruby

### 2012-01-03 paper.tex
MOAR text and cleaner first page

@@

\section{Comparisons}

@@

@@

class Cat < Animal
	def speak
		puts "Meow"
	end
end
\end{lstlisting}
When using these two classes, one can see that methods will be overriden:
\begin{lstlisting}[language=Ruby]
cat = Cat.new
cat.speak
cat.breath
\end{lstlisting}
This will output:
\begin{lstlisting}
Random noise
Breathe
\end{lstlisting}
Inheritance in C++ is more advanced as it supports multiple inheritance.

Ruby mixin's.

@@

\subsection{The use of super and this/self}

>>>>

\tableofcontents

\columnbreak

@@

\pagebreak
\section{Comparisons}

@@

\subsubsection{Friend classes}
C++ has an unique keyword: \textit{friend}. With this, one can declare a class as a friend:

\begin{lstlisting}
class A;
class B {
	friend class A;
}
\end{lstlisting}

In this example, class A is a friend of B. This means that class A has more access rights to the members and methods of class B. More concrete, this means that all private and protected members and methods of B are accessible by A while this would not be the case by default. The \textit{friend} keyword defines the following features:

\begin{itemize}
\item Friendships are not corresponded: if class A is a friend of class B, class B is not automatically a friend of class A.
\item Friendships are not transitive: if class A is a friend of class B and class B is a friend of class C, then is class A not a friend of class C.
\item Friendships are not inherited: if there is a class Base with a friend, then the class Derived (that inherits from Base) does not have that same friend. The same goes up for all friends of Derived, they are not automatically friends of Base. The same goes in the other direction: if Base (or Derived) is a friend of a class, then is Derived (or Base) not automatically a friend of the same class.
\item The restricted members and methods from Base that are inherited by Derived are accessible by a friend class of Derived if Derived itself can access these members and methods. So if Derived inherits publicly from Base, Derived only has access to the protected and public members from Base, not the private members, so neither does a friend (the same goes for methods).
\end{itemize}

@@

class Cat < Animal
	def speak
		puts "Meow"
	end
end
\end{lstlisting}
When using these two classes, one can see that methods will be overriden:
\begin{lstlisting}[language=Ruby]
cat = Cat.new
cat.speak
cat.breath
\end{lstlisting}
This will output:
\begin{lstlisting}
Random noise
Breathe
\end{lstlisting}

C++ supports multiple inheritance, with this it is possible to inherit from multiple classes like this:

\begin{lstlisting}
class C: public A, protected B { };
\end{lstlisting}

In this example, a class C is created that inherits publicly from class A and protectedly from class B. As you can see, the access modifiers can be different. One of the common pitfalls when using multiple inheritance in C++ is ambiguities:

\begin{lstlisting}
class A { virtual void f(); };
class B { virtual void f(); };
class C : public A ,public B { void f(); };
\end{lstlisting}

This issue can be solved by using explicit qualification. We explicitly tell the compiler which \textit{f()} we would like to call:

\begin{lstlisting}
C* c = new C;
c->f();
c->A::f(); // call f() from class A
c->B::f(); // call f() from class B
\end{lstlisting}

Another more complex problem which can arise with multiple inheritance is called the \textit{diamond problem}. The diamond problem is an ambiguity that arises when two classes B and C inherit from A, and class D inherits from both B and C. If a method in D calls a method that is defined in A (and D does not override the method) and B and C have overridden that method differently, then from which class does it inherit: B or C? This is a problem that can be solved by using virtual inheritance.

When constructing an object of class D, C++ actually creates in the memory objects of B and C internally. B and C in their turn also construct each an object of A. This means that upon the creation of D, 4 internal objects are created: B, C and 2 times A. If we however have inherited B and C from A as virtual (for example: \textit{class B: virtual public A}), C++ will make sure that only one object of A is constructed.

Although Ruby does not support multiple inheritance, it eliminates the need for multiple inheritance by providing mixins. As a programmer, you can define modules in Ruby. In classes, you can then include these modules. For example:

\begin{lstlisting}[language=Ruby]
module Foo
	def Bar
		puts "Wazaa!"
	end
end

class FooBar
	include Foo
end
\end{lstlisting}

In this example, the class FooBar contains the content of module Foo (in this case the method Bar). It is possible to include multiple modules.

@@
